Fixed-Parameter Tractable Sampling for RNA Design with Multiple Target Structures

### 2019-01-24 supplement.tex
Updates from ShareLaTeX

For each of two small examples, a single-stem and a double-stem structure, we produce a uniform and a weighted Boltzmann
sampling, each of 10000 sequenes. For each sampled sequence $S$, we determine
the Turner energy $E(S,R)$ as well as the
%difference $E_\Delta(S,R)$ of $E(S,R)$ to the ensemble energy $E_{\text{ens}}(S)$ of $S$ 
probability $Pr(R|S)$ of the structure $R$ in the ensemble of sequence $S$ 
(where the energies are calculated in the Turner model using RNAfold).

@@

To illustrate the efficiency and practical feasibility of our Boltzmann sampling approach, we show the distributions of run times (Intel(R) Core(TM) i7-4770 CPU) for generating 10000 samples (over the Modena benchmark instances and grouped by benchmark sub-set). Since producing a controlled Boltzmann sampling requires significant pre-computation, it is informative to separately show construction and sampling times (Figure~\ref{appfig:run-times-redprint}). Moreover, \ourprog{} can be used with a simpler base pair model or a rather complex stacking energy model, which introduces much stronger dependencies than required for uniform sampling. Technically, these stronger dependencies increase the tree width and thus the runtime, as noticeable in Figure~\ref{appfig:run-times-redprint}.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/construction_time_sets}\quad%
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/sample_time_sets}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \ourprog{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the pre-computation, while the latter is the time needed to generate 10000 sequences. }
  \label{appfig:run-times-redprint}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/construction_time_sets_rbp}\quad%
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/sample_time_sets_rbp}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \RNAblueprint{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the pre-computation, while the latter is the time needed to generate 10000 sequences.}
  \label{appfig:run-times-rbp}
\end{figure}

@@

In analogy to Fig.~\ref{fig:energydist-pk} of the main text,
    Fig.~\ref{appfig:energydist} illustrates our multi-dimensional sampling
    strategy for three pseudoknot-free target structures (example instance
    from the \texttt{Modena} benchmark). The figure shows Turner energy
    distributions of single structures from sampling with different weight
    parameters, where weights are trained to target specific energies. For
    comparison, the figure shows distributions from uniform and Boltzmann
    sampling at respective homogenous weights $1$ and $e^\beta$.

@@

\begin{figure}[h!]
      \begin{center}
        %{\sf \bfseries A}\includegraphicstop[width=0.32\textwidth]{Figs/offset2}\hspace{.5cm}
        %{\sf \bfseries B}
        \includegraphicstop[width=0.8\textwidth]{Figs/energy_distribution}
      \end{center}
      \caption{%
        Targeting specific energies using multi-dimensional Boltzmann
        sampling. We show the Turner energy distributions for
        three target structures R1, R2 and R3. By
        targeting the respective free energies $(-40,-40,-20)$ and
        $(-20,-20,-20)$ for the target structures (annotated in red and
        blue, resepectively), we demonstrate the
        effectivity of our adaptive multi-dimensional Boltzmann sampling
        procedure; moreover, for comparison, the distributions for uniform
        (U) and Boltzmann (B) samples---respectively associated with
        homogenous weights $1$ and $e^\beta$.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
  }
  \label{appfig:energydist}
\end{figure}

@@

    %vertex caracterizes the parity of all vertices in a bipartite

>>>>

For each of two small examples, a single-stem and a double-stem structure, we produce a uniform and a weighted Boltzmann
sampling, each of 10\,000 sequences. For each sampled sequence $S$, we determine
the Turner energy $E(S,R)$ as well as the
%difference $E_\Delta(S,R)$ of $E(S,R)$ to the ensemble energy $E_{\text{ens}}(S)$ of $S$ 
probability $Pr(R|S)$ of the structure $R$ in the ensemble of sequence $S$ 
(where the energies are calculated in the Turner model using RNAfold).

@@

To illustrate the efficiency and practical feasibility of our Boltzmann sampling approach, we show the distributions of run times (Intel(R) Core(TM) i7-4770 CPU) for generating 10000 samples (over the Modena benchmark instances and grouped by benchmark sub-set). Since producing a controlled Boltzmann sampling requires significant precomputation, it is informative to separately show construction and sampling times (Figure~\ref{appfig:run-times-redprint}). Moreover, \ourprog{} can be used with a simpler base pair model or a rather complex stacking energy model, which introduces much stronger dependencies than required for uniform sampling. Technically, these stronger dependencies increase the tree width and thus the run-time, as noticeable in Figure~\ref{appfig:run-times-redprint}.

\begin{figure}[h!]
  \centering
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/construction_time_sets}\quad%
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/sample_time_sets}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \ourprog{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the precomputation, while the latter is the time needed to generate 10000 sequences. }
  \label{appfig:run-times-redprint}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/construction_time_sets_rbp}\quad%
  \includegraphics[width=0.43\textwidth]{Figs/RunTimes/sample_time_sets_rbp}
  \caption{Boxplots showing the run-time distribution of sequence sampling with \RNAblueprint{} for the various benchmark sets. The times are split into construction time (left) and sampling time (right). Construction includes all the precomputation, while the latter is the time needed to generate 10000 sequences.}
  \label{appfig:run-times-rbp}
\end{figure}

@@

In analogy to Fig.~\ref{fig:energydist-pk} of the main text,
    Fig.~\ref{appfig:energydist} illustrates our multi-dimensional sampling
    strategy for three pseudoknot-free target structures (example instance
    from the \texttt{Modena} benchmark). The figure shows Turner energy
    distributions of single structures from sampling with different weight
    parameters, where weights are trained to target specific energies. For
    comparison, the figure shows distributions from uniform and Boltzmann
    sampling at respective homogeneous weights $1$ and $e^\beta$.

@@

\begin{figure}[h!]
      \begin{center}
        %{\sf \bfseries A}\includegraphicstop[width=0.32\textwidth]{Figs/offset2}\hspace{.5cm}
        %{\sf \bfseries B}
        \includegraphicstop[width=0.8\textwidth]{Figs/energy_distribution}
      \end{center}
      \caption{%
        Targeting specific energies using multi-dimensional Boltzmann
        sampling. We show the Turner energy distributions for
        three target structures R1, R2 and R3. By
        targeting the respective free energies $(-40,-40,-20)$ and
        $(-20,-20,-20)$ for the target structures (annotated in red and
        blue, respectively), we demonstrate the
        effectivity of our adaptive multi-dimensional Boltzmann sampling
        procedure; moreover, for comparison, the distributions for uniform
        (U) and Boltzmann (B) samples---respectively associated with
        homogeneous weights $1$ and $e^\beta$.
%    ($\delta=0.05, \delta'=0.1, \gamma=1.1, \psi=20, n=1000$)
  }
  \label{appfig:energydist}
\end{figure}

@@

    %vertex characterizes the parity of all vertices in a bipartite
